```{r}
library(tidyverse)
library(anndata)
library(devtools)
library(ComplexHeatmap)
library(circlize)
library(RColorBrewer)
library(gridExtra)
library(forcats)
library(ggplot2)
library(survival)
library(survminer)
```
```{r}
#set heatmap main coloring
main_col_fun = colorRamp2(
  c(-1.5,0,1.5), 
  c("#2166ac", "grey", "#b2182b"))

main_col_fun = colorRamp2(
  c(-1.5,0,1.5), 
  c("#2166ac", "#f7f7f7", "#b2182b"))

```

```{r}

dir <- "~/Desktop/PhD/NMC/results"

infiltrated_cells <- read.csv(file.path(dir, "NMC_cozi_infiltration.csv")) %>%
  mutate("infiltration" = cond_cells_percentage) %>%
  select(!cond_cells_percentage)

#read interaction file
interaction_immune <- read.csv(file.path(dir, "NMC_immunetumor_interaction.csv"))

interaction_tumor <- read.csv(file.path(dir, "NMC_tumorimmune_interaction.csv"))

infiltration_percentage <- read.csv(file.path(dir, "NMC_infiltration_percentage.csv"))

interaction_MDSCs <- read.csv(file.path(dir, "NMC_MDSCsimmune_interaction.csv"))

```


```{r}
#Build with complex heatmaps two heatmaps 

interaction_immune_mat <- interaction_immune %>%
  pivot_wider(names_from = "patient_exp", values_from = "zscore") %>%
  column_to_rownames(var = "cell_type") %>%
  as.matrix() %>% 
  scale() %>%
  t()

interaction_tumor_mat <- interaction_tumor %>%
  pivot_wider(names_from = "patient_exp", values_from = "zscore") %>%
  column_to_rownames(var = "cell_type") %>%
  as.matrix() %>%
  scale() %>%
  t()

interaction_MDSCs_mat <- interaction_MDSCs %>%
  pivot_wider(names_from = "patient_exp", values_from = "zscore") %>%
  column_to_rownames(var = "cell_type") %>%
  as.matrix() %>%
  scale() %>%
  t()

Heatmap(interaction_immune_mat,
        name = "Immune to Tumor Interaction (z_score)",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = TRUE,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(interaction_immune_mat[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)

#do same heatmap for tumor interaction

Heatmap(interaction_tumor_mat,
        name = "Tumor to Immune Interaction (z_score)",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = TRUE,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(interaction_tumor_mat[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)


Heatmap(interaction_MDSCs_mat,
        name = "Immune to Tumor Interaction (z_score)",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = TRUE,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(interaction_MDSCs_mat[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)

```
```{r}
#Let's see if MDSCs interaction is correlated negatively with CD8+_T_cells

interaction_cor = interaction_immune %>%
  group_by(patient_exp) %>%
  mutate("zscore" = scale(zscore)) %>%
  #make interactions na 0
  mutate(zscore = ifelse(is.na(zscore), 0, zscore)) %>%
  ungroup() %>%
  #fill NA with 0
  filter(cell_type %in% c("MDSCs", "CD8+_T_cells", "M1_M2_macrophages","NK_cells", "M1_macrophages", "M2_macrophages"))

#let's do a ggplot as boxplot to compare
ggplot(interaction_cor, aes(x = cell_type, y = zscore, fill = cell_type)) +
  geom_boxplot() +
  geom_jitter() + 
  labs(title = "Interaction between MDSCs and CD8+ T cells",
       x = "Cell Type",
       y = "Z-score") +
  theme_minimal() +
  stat_compare_means(method = "t.test", label = "p.format", label.y = 2.5)


#let's do a ggplot with linear regression
#pivot the data wider for it first
interaction_cor_wide <- interaction_cor %>%
  pivot_wider(names_from = "cell_type", values_from = "zscore")
ggplot(interaction_cor_wide, aes(x = MDSCs, y = `CD8+_T_cells`)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Correlation between MDSCs and CD8+ T cells",
       x = "MDSCs Interaction (z-score)",
       y = "CD8+ T cells Interaction (z-score)") +
  theme_minimal() +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top")

```
```{r}

interaction_cor = interaction_immune %>%
  group_by(patient_exp) %>%
  mutate("zscore" = scale(zscore)) %>%
  #make interactions na 0
  mutate(zscore = ifelse(is.na(zscore), 0, zscore)) %>%
  ungroup()

# for loop to get correlation between all cell types

cell_types <- c("MDSCs", "M1_macrophages", "CD8+_T_cells", "M2_macrophages", "NK_cells", "M1_M2_macrophages")

interaction_cor <- interaction_cor %>%
  filter(cell_type %in% cell_types) %>%
  mutate("group" = ifelse(cell_type %in% c("MDSCs", "M2_macrophages", "M1_M2_macrophages"), "immune_supressive", "immune_effector"))

ggplot(interaction_cor, aes(x = group, y = zscore, fill = group)) +
  geom_boxplot() +
  geom_jitter() + 
  labs(title = "Interaction Correlation between Cell Types",
       x = "Cell Type",
       y = "Z-score") +
  theme_minimal() +
  stat_compare_means(method = "t.test", label = "p.format", label.y = 2.5)

```


interaction_cor_df = data.frame()

for(i in cell_types) {
  for(j in cell_types) {
    if(i == j){
      cor_value = NA
      interaction_cor_df <- rbind(interaction_cor_df, data.frame(cell_type1 = i, cell_type2 = j, correlation = cor_value))
    } else{
    cor_value <- cor(interaction_cor %>% filter(cell_type == i) %>% pull(zscore),
                       interaction_cor %>% filter(cell_type == j) %>% pull(zscore), use = "complete.obs")
      interaction_cor_df <- rbind(interaction_cor_df, data.frame(cell_type1 = i, cell_type2 = j, correlation = cor_value))
    }
  }
}


#Build a heatmap with the correlation values
interaction_cor_matrix <- interaction_cor_df %>%
  pivot_wider(names_from = cell_type2, values_from = correlation) %>%
  column_to_rownames(var = "cell_type1") %>%
  as.matrix()

main_col_fun = colorRamp2(
  c(-0.5,0,0.5), 
  c("#2166ac", "#f7f7f7", "#b2182b"))

Heatmap(interaction_cor_matrix,
        name = "Interaction Correlations",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "Correlation"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = FALSE,
        column_title = "Cell Types",
        row_title = "Cell Types",
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(interaction_cor_matrix[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)

```


```{r}

infiltration_percentage_cor <- infiltration_percentage %>%
  column_to_rownames(var = "cell_type") %>%
  t() %>%
  as.data.frame() %>%
  select("MDSCs", "CD8+_T_cells") %>%
  pivot_longer(names_to = "cell_type", values_to = "infiltration", cols = everything()) 

infiltrated_cells_cor <- infiltrated_cells %>%
  filter(cell_type %in% c("MDSCs", "CD8+_T_cells"))

#make boxplot to compare 
ggplot(infiltrated_cells_cor, aes(x = cell_type, y = infiltration, fill = cell_type)) +
  geom_boxplot() +
  geom_jitter() + 
  labs(title = "Infiltration Percentage of MDSCs and CD8+ T cells into the Tumor",
       x = "Cell Type",
       y = "Infiltration Percentage") +
  theme_minimal() +
  stat_compare_means(method = "t.test", label = "p.format", label.y = 2.5)
```


```{r}

# Convert to wide matrices
zscore_mat <- interaction %>%
  pivot_wider(names_from = "patient_exp", values_from = "zscore") %>%
  column_to_rownames(var = "cell_type") %>%
  as.matrix() %>% 
  scale() %>%
  t()

infiltration_mat <- infiltrated_cells %>%
  pivot_wider(names_from = "patient_exp", values_from = "infiltration") %>%
  column_to_rownames("cell_type") %>%
  as.matrix() %>%
  t()


main_col_fun = colorRamp2(
  c(-1.5,0,1.5), 
  c("#2166ac", "grey", "#b2182b"))

cell_drawer <- function(j, i, x, y, width, height, fill) {
  score <- zscore_mat[i, j]
  infiltration <- infiltration_mat[i, j]
  
  if (!is.na(score) && !is.na(infiltration)) {

    # Bin into 4 levels
    radius_factor <- case_when(
      infiltration <= 0.2 ~ 0.25,
      infiltration <= 0.4  ~ 0.5,
      infiltration <= 0.6  ~ 0.75,
      TRUE               ~ 1.0
    )
    
    max_radius <- unit(0.035, "snpc")
    radius <- radius_factor * convertUnit(max_radius, "npc", valueOnly = TRUE)

    grid.circle(
      x = x, y = y,
      r = unit(radius, "npc"),
      gp = gpar(fill = main_col_fun(score), col = NA)
    )
  } else {
    scale_factor <- 0.3  
    w2 <- width * scale_factor
    h2 <- height * scale_factor
    grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 2))
    grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 2))
    
  }
}

jpeg(filename = "Desktop/heatmap_output.jpeg", width = 1200, height = 800, res = 150) 

ht = Heatmap(
  matrix = zscore_mat,
  name = "Interaction",
  cell_fun = cell_drawer,
  cluster_rows = FALSE,
  cluster_columns = TRUE,
  column_names_rot = 45,
  row_names_side = "left",
  rect_gp = gpar(col = "black", fill = NA),
  col = main_col_fun,
  column_title = "Cell Types",
  row_title = "Samples",
  column_names_gp = gpar(fontsize = 8),
  row_names_gp = gpar(fontsize = 8),
)

circle_legend <- Legend(
  title = "Infiltration",
  type = "points",
  pch = 21,
  size = unit(c(0.15, 0.30, 0.45, 0.6), "cm"),
  labels = c("0.2", "0.4", "0.6", "0.8"),
  title_gp = gpar(fontsize = 10, fontface = "bold"),
  labels_gp = gpar(fontsize = 10),
  grid_height = unit(0.55, "cm"),
  grid_width = unit(0.7, "cm")
)

# Draw heatmap and legend together
draw(ht, annotation_legend_list = list(circle_legend))

dev.off()

test = infiltrated_cells %>% 
  group_by(cell_type) %>%
  summarise("median_infiltration" = mean(infiltration, na.rm = TRUE))
```


```{r}

# Convert to wide matrices
interaction_boxplot <- interaction %>%
  group_by(patient_exp) %>%
  mutate(zscore = scale(zscore)) %>%
  ungroup() %>%
  filter(cell_type %in% c("MDSCs", "M2_macrophages", "Granulocytes", "CD4+_T_cells", "CD8+_T_cells"))

boxplot_plot = ggplot(interaction_boxplot, aes(x = cell_type, y = zscore, fill = cell_type)) +
  geom_boxplot() +
  geom_jitter() + 
  stat_compare_means()

ggsave("~/Desktop/boxplot.jpeg", plot = boxplot_plot, device = "jpeg")

```

```{r}
Heatmap(mat_scaled,
        name = "distances_to_tumor_core",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction (z_score)"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = FALSE,
        top_annotation = column_ha,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        right_annotation = row_ha,
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(mat_scaled[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)

```






```{r}
interaction = read.csv(file.path(dir, "NMC_immunetumor_interaction.csv")) %>%
  na.omit() %>%
  filter(!cell_type == "Epithelial_cells") %>% 
  mutate("patient_ID" = str_split_fixed(patient_exp, "_ROI", 2)[,1]) %>%
  arrange(patient_exp, factor(cell_type, levels = infiltrated_cells$cell_type))

#survival time

survival = readxl::read_xlsx("Desktop/PhD/NMC/Metadata/extra_patients_info.xlsx", sheet = "Detailed info") %>%
  rename("patient_ID" = "Patient ID") %>%
  select(patient_ID, `Survival after NMC diagnosis`) %>%
  inner_join(distances[,c("patient_exp", "patient_ID")]) %>%
  arrange(patient_exp) %>% 
  unique() %>%
  rename("survival_time" = `Survival after NMC diagnosis`) %>%
  mutate(survival_months = str_extract(survival_time, "\\d+")) %>%
  mutate(survival_months = as.numeric(survival_months)) %>%
  arrange(survival_months)

distances = distances %>%
   arrange(factor(patient_exp, survival$patient_exp), factor(cell_type, levels = infiltrated_cells$cell_type)) 

survival_vector = survival %>% 
  pull(survival_months)

heatmap_distances = distances %>%
  select(!patient_ID) %>%
  pivot_wider(names_from = "patient_exp", values_from = "median_distance") %>%
  column_to_rownames(var = "cell_type")

mat <- as.matrix(heatmap_distances)

# scale by row (cell type)
mat_scaled <- scale(mat)
mat_scaled = t(mat_scaled)


infiltration = infiltrated_cells$infiltrated_percentage

main_col_fun = colorRamp2(
  c(-1.5,0,1.5), 
  c("#2166ac", "#f7f7f7", "#b2182b")
)
# This gives nice dark blue for negative z, white for zero, dark red for high z

# Define color scale for infiltration percentage annotation (green)
annotation_col_fun = colorRamp2(
  c(0, 30, 60),
  c("#f7fcf5", "#74c476", "#00441b")
)

# Build column annotation
column_ha = HeatmapAnnotation(
  "Infiltration (%)" = infiltration,
  col = list("Infiltration (%)" = annotation_col_fun),
  annotation_name_gp = gpar(fontsize = 10)
)

survival_bins <- cut(survival_vector, 
                     breaks = c(0, 6, 12, Inf), 
                     labels = c("<6m", "6-24m", ">24m"),
                     right = FALSE)

# Assign colors to bins
bin_colors <- c("<6m" = "#d73027", 
                "6-24m" = "#fc8d59", 
                ">24m" = "#1a9850")

# Create annotation
row_ha <- rowAnnotation("Survival group" = survival_bins,
                        col = list("Survival group" = bin_colors),
                        annotation_name_gp = gpar(fontsize = 10))


jpeg(filename = "Desktop/heatmap_output.jpeg", width = 1200, height = 800, res = 150) 

Heatmap(mat_scaled,
        name = "distances_to_tumor_core",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction (z_score)"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = FALSE,
        top_annotation = column_ha,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        right_annotation = row_ha,
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(mat_scaled[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)
dev.off()


```

```{r}
interaction_patient <- distances %>%
  group_by(patient_ID, cell_type) %>%
  summarise(interaction_score = median(median_distance, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = cell_type, values_from = interaction_score)

# 2️⃣ Merge with survival data

surv_data <- survival %>%
  select(patient_ID, survival_months) %>%
  distinct()  # remove duplicate rows due to multiple ROIs

full_data <- inner_join(interaction_patient, surv_data, by = "patient_ID")

full_data <- full_data %>%
  mutate(
    time = survival_months,
    event = 1  # assuming all events are observed; update if you have censoring
  )


results <- list()

for(cell in setdiff(colnames(full_data), c("patient_ID", "survival_months", "time", "event"))) {
  df <- full_data %>% select(time, event, all_of(cell)) %>% drop_na()
  
  if(nrow(df) >= 5) {  # require minimal number of patients for model stability
    model <- coxph(Surv(time, event) ~ ., data = df)
    summary_model <- summary(model)
    
    results[[cell]] <- data.frame(
      cell_type = cell,
      HR = summary_model$coef[1, "exp(coef)"],
      p_value = summary_model$coef[1, "Pr(>|z|)"],
      n = nrow(df)
    )
  }
}

results_df <- do.call(rbind, results)

results_df$cell_type <- factor(results_df$cell_type, levels = results_df$cell_type[order(results_df$p_value)])

ggplot(results_df, aes(x = HR, y = cell_type)) +
  geom_point() +
  geom_errorbarh(aes(xmin = HR * exp(-1.96 * sqrt(1/n)), 
                     xmax = HR * exp(1.96 * sqrt(1/n))), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  xlab("Hazard Ratio") + ylab("Cell Type") +
  theme_minimal() +
  ggtitle("Univariate Cox Hazard Ratios")
```
```{r}

distances_compare = distances %>%
  left_join(survival) %>%
  mutate("survival_group" = ifelse(
    survival_months > 6,
    "Long",
    "Short"
  )) %>%
  filter(cell_type %in% c("MDSCs", "Granulocytes", "M2_macrophages", "B_cells")) %>%
  group_by(cell_type) %>%
  mutate(median_distance = scale(median_distance)) %>%
  select(patient_exp, cell_type, median_distance, survival_group)
  

ggplot(distances_compare, aes(x = survival_group, y = median_distance, fill = survival_group)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  geom_jitter(width = 0.15, size = 2, alpha = 0.8) +
  facet_wrap(~ cell_type, scales = "free_y") +
  scale_fill_manual(values = c("Short" = "#F8766D", "Long" = "#00BFC4")) +
  labs(
    title = "Interaction Scores by Survival Group",
    subtitle = "MDSCs, Granulocytes, and B cells",
    x = "Survival Group",
    y = "Scaled Interaction Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 13)
  ) +
   stat_compare_means(
    method = "wilcox.test",   # safer for small n
    label = "p.format",       # nicely formatted p-value
    comparisons = NULL,       # auto-detect comparisons
    label.y = NULL            # auto-place p-values
  )

```

```{r}

infiltration_new = read.csv("Desktop/infiltration_percentage.csv") %>%
  arrange(factor(cell_type, levels = rownames(mat))) %>%
  column_to_rownames(var = "cell_type") %>%
  select(colnames(mat)) %>% 
  as.matrix()

h1 = Heatmap(mat_scaled,
        name = "distances_to_tumor_core",
        column_names_rot = 45,
        heatmap_legend_param = list(title = "interaction (z_score)"),
        cluster_rows  = FALSE,
        col = main_col_fun,
        cluster_columns = FALSE,
        top_annotation = column_ha,
        column_title = "Cell Types",
        row_title = "Samples",
        row_names_side = "left",
        right_annotation = row_ha,
        column_names_gp = gpar(fontsize = 8),
        row_names_gp = gpar(fontsize = 8),
        na_col = "white",  ## background color for NA
        cell_fun = function(j, i, x, y, width, height, fill) {
            if (is.na(mat_scaled[i, j])) {
                # Draw smaller X
                scale_factor <- 0.3  
                w2 <- width * scale_factor
                h2 <- height * scale_factor
                grid.lines(x = c(x - w2, x + w2), y = c(y - h2, y + h2), gp = gpar(col = "black", lwd = 1.5))
                grid.lines(x = c(x - w2, x + w2), y = c(y + h2, y - h2), gp = gpar(col = "black", lwd = 1.5))
            }
        }
)

h2 = Heatmap(t(infiltration_new), 
             column_names_rot = 45,
             cluster_rows  = FALSE,
             cluster_columns = FALSE,
             column_title = "Cell Types",
             row_title = "Samples",
             row_names_side = "left",
             column_names_gp = gpar(fontsize = 8),
             row_names_gp = gpar(fontsize = 8),
             na_col = "white")


h1 + h2
```

```{r}

cor = distances %>% 
  filter(cell_type %in% c("MDSCs", "CD8+_T_cells")) %>%
  pivot_wider(names_from = "cell_type", values_from = "median_distance")


ggplot(cor, aes(x=`MDSCs`, y=`CD8+_T_cells`)) +
  geom_point() +
  geom_smooth(method = "lm")


```

```{r}

counts_df = read.csv("Desktop/cell_category_counts.csv") %>%
  select(!cell_type) %>%
  rename(Tissue = site.of.biopsy.resection.for.MACSima.sample) 


counts_exp <- counts_df %>% 
  group_by(patient_exp, cell_category) %>%
  summarise(Frequency = n()) %>% 
  mutate(Percentage = 100 * Frequency / sum(Frequency)) %>%
  ungroup()

counts_patient <- counts_df %>% 
  group_by(patient_ID, cell_category) %>%
  summarise(Frequency = n()) %>% 
  mutate(Percentage = 100 * Frequency / sum(Frequency)) %>%
  ungroup()

counts_tissue <- counts_df %>% 
  group_by(Tissue, cell_category) %>%
  summarise(Frequency = n()) %>% 
  mutate(Percentage = 100 * Frequency / sum(Frequency)) %>%
  ungroup()

counts_exp$cell_category %>% unique()
```


```{r}

# Define cell type colors
cell_category_color <- c(
  "Actin+ cells" = "#AAAAFF",
  "B cells" = "#FFFF00",
  "DCs" = "#00FF00",
  "Endothelial cells" = "#55FFFF",
  "Granulocytes" = "#FF00FF",
  "Lymphatic endothelial cells" = "#0055FF",
  "Macrophages" = "#00557F",
  "Mast cells" = "#AA007F",
  "MDSCs" = "#00007F",
  "NFC" = "#A9A9A9",
  "NK cells" = "#00AAFF",
  "PCs" = "#FFFF7F",
  "T cells" = "#FF0000FF",
  "Tumor cells" = "#F8F8F8",
  "Epithelial cells" = "#FFCC99"
)

cells_category_order <- c(
  'Actin+ cells',
  'Endothelial cells',
  'Lymphatic endothelial cells',
  'Macrophages',
  'Granulocytes',
  'MDSCs',
  'Mast cells',
  'DCs',
  'NK cells',
  'T cells',
  'B cells',
  'PCs',
  'Epithelial cells',
  'NFC',
  'Tumor cells'
)

# Assume `adata.obs` has been imported into R as a dataframe `adata_obs`
# and has columns: patient_exp, cell_category, patient_ID, site

# Function to compute % composition
compute_composition <- function(df, group_var) {
  df %>%
    group_by(across(all_of(group_var)), cell_category) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(across(all_of(group_var))) %>%
    mutate(pct = round(n / sum(n) * 100, 2)) %>%
    ungroup()
}

# Apply to each grouping
counts_exp <- compute_composition(counts_df, "patient_exp")
counts_patient <- compute_composition(counts_df, "patient_ID")
counts_tissue <- compute_composition(counts_df, "Tissue")

# Function to create a plot
make_bar_plot <- function(data, x_var, title) {
  ggplot(data, aes_string(x = x_var, y = "pct", fill = "cell_category")) +
    geom_bar(stat = "identity", position = "stack") +
    scale_fill_manual(values = cell_category_color, breaks = cells_category_order) +
    labs(x = x_var, y = "Percentage of cells", title = title, fill = "Cell type") +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
      plot.title = element_text(hjust = 0.5)
    )
}

# Generate each plot
p1 <- make_bar_plot(counts_exp, "patient_exp", "Cell type proportions per ROI") +
  theme(legend.position = "none")
p2 <- make_bar_plot(counts_patient, "patient_ID", "Cell type proportions per patient") +
  theme(legend.position = "none")
p3 <- make_bar_plot(counts_tissue, "Tissue", "Cell type proportions per tissue")

# Combine plots
(p1 | p2 | p3) + plot_layout(guides = "collect") & theme(legend.position = "right")



```

```{r}
cell_category_color = c("Actin+ cells" = "#AAAAFF",
                        "B cells" = "#FFFF00",
                        "Dendritic cells" = "#00FF00",
                        "Endothelial cells" = "#55FFFF",
                        "Granulocytes" = "#FF00FF",
                        "Lymphatic endothelial cells" = "#0055FF",
                        "Macrophages" = "#00557F",
                        "Mast cells" = "#AA007F",
                        "Myeloid-derived suppressor cells (MDSCs)" = "#00007F",
                        "NFC (Not Further Characterized)" = "#550000",
                        "NK cells" = "#00AAFF",
                        "Plasma cells" = "#FFFF7F",
                        "T cells" = "#FF0000",
                        "Tumor cells" = "#8B0000",
                        "Epithelial cells" = "#FF8C00"
                        )
  
ggplot(counts_exp, aes(x = patient_exp, y = Percentage, fill = cell_category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = cell_category_color) +
  labs(title = "Cell Type Proportions per ROI", x = "ROI", y = "Proportion") 


```

```{r}
tumor_prop = read.csv("Desktop/tumor_markers_percentage.csv")

tumor_markers = c("CD276", "CD326", "CLEAVEDPARP1", "CYTOKERATIN14", "CYTOKERATIN19",
  "CYTOKERATIN5", "CYTOKERATIN7", "ERBB2", "KI67", "P16", "P53",
  "PCNA", "PRAME")

tumor_prop = tumor_prop %>%
  mutate("tumor_percentage" = case_when(tumor_cells_percentage <= 25 ~ "0-25%",
                                        tumor_cells_percentage > 25 & tumor_cells_percentage < 50 ~ "25-50%",
                                        tumor_cells_percentage >= 50 & tumor_cells_percentage < 75 ~ "50-75%",
                                        tumor_cells_percentage >= 75 ~ "75-100%"))

colors <- c("0-25%" = "#d73027",   # red
            "25-50%" = "#fc8d59",  # orange
            "50-75%" = "#fee08b",  # yellow
            "75-100%" = "#1a9850")    # green



tumor_prop_piv = tumor_prop %>%
  select(patient_id, marker, tumor_cells_percentage) %>%
  pivot_wider(names_from = "patient_id", values_from = "tumor_cells_percentage") %>%
  arrange(marker) %>%
  column_to_rownames(var = "marker") %>%
  select(!NMC_9)

tumor_prop_piv = tumor_prop_piv 

tumor_count = read.csv("Desktop/tumor_count.csv") %>%
  filter(!patient_ID == "NMC_9")
```

```{r}
tumor_count_ordered <- tumor_count[match(colnames(tumor_prop_piv), tumor_count$patient_ID), ]

column_ha = HeatmapAnnotation("% of cells in the sample" = anno_barplot(tumor_count_ordered$Tumor_cells))

Heatmap(tumor_prop_piv,cluster_columns = FALSE, 
        rect_gp = gpar(col = "black", lwd = 1), 
        cluster_rows = FALSE, 
        column_title = "Percentage of Marker-Positive Tumor Cells Per Patient",
        heatmap_legend_param = list(title = "Percentage"),
        top_annotation = column_ha,
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.1f%%", tumor_prop_piv[i, j]), x, y, gp = gpar(fontsize = 8))
        })

column_ha = HeatmapAnnotation("% of cells in the sample" = anno_barplot(tumor_count_ordered$Tumor_cells))

Heatmap(tumor_prop_piv,cluster_columns = FALSE, 
        rect_gp = gpar(col = "black", lwd = 1), 
        cluster_rows = FALSE, 
        column_title = "Percentage of Marker-Positive Tumor Cells Per Patient",
        heatmap_legend_param = list(title = "Percentage"),
        top_annotation = column_ha)


```


```{r}
#read df pos
df_pos = read.csv("Desktop/df_pos_full.csv")
df_pos
m = make_comb_mat(df_pos)

pdf("Desktop/upset_wide.pdf", width = 30, height = 10)  # Width in inches
UpSet(m)
dev.off()
```

```{r}
distances_df = read.csv("Desktop/distance_df.csv")
distances_df = distances_df %>%
  select(!c(patient_ID, distance_to)) %>%
  group_by(cell_type, patient_exp) %>%
  summarise(mean_distance = mean(distance, na.rm = TRUE))
distances_df$mean_distance = as.numeric(distances_df$mean_distance)

distances_df = distances_df %>%
  pivot_wider(names_from = "patient_exp", values_from = "mean_distance")

mat <- distances_df %>%
  column_to_rownames("cell_type") %>%
  as.matrix() %>%
  apply(2, as.numeric)  # Ensure all columns are numeric

rownames(mat) = distances_df$cell_type

pdf("Desktop/distances_df.pdf", width = 15, height = 10)  # Width in inches
Heatmap(mat,cluster_columns = FALSE, 
        rect_gp = gpar(col = "black", lwd = 2), 
        cluster_rows = FALSE, 
        column_title = "Mean distance between Cell type of interest and tumor cells",
        heatmap_legend_param = list(title = "Distance"),
  )
dev.off()
```

```{r}
heatmap_df = read.csv("Desktop/heatmap_data.csv")

mat <- heatmap_df %>%
  column_to_rownames("X") %>%
  as.matrix() %>%
  apply(2, as.numeric) 

rownames(mat) = heatmap_df$X

pdf("Desktop/heatmap_df.pdf", width = 15, height = 10)  # Width in inches
Heatmap(mat,cluster_columns = FALSE, 
        rect_gp = gpar(col = "black", lwd = 2), 
        cluster_rows = FALSE, 
        column_title = "Mean distance between Cell type of interest and tumor cells",
        heatmap_legend_param = list(title = "Distance"),
        cell_fun = function(j, i, x, y, width, height, fill) {
          grid.text(sprintf("%.1f", mat[i, j]), x, y, gp = gpar(fontsize = 10, col = "white"))
        }
    )
dev.off()


```


```{r}
boxplot = read.csv("Desktop/boxplot_data_select.csv")

ggplot(boxplot, aes(x = patient_ID, y = zscore, fill = interaction)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.2, size = 2, alpha = 0.7, color = "darkblue") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Z-scores per Interaction",
    x = "Cell-Cell Interaction",
    y = "Z-score"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```